---
title: "Precision simulations"
author: "Ruben Arslan"
date: "2023-11-07"
output: 
  html_document:
    toc: true
    toc_float: true
---


The following precision simulations all follow the same structure.

- We randomly redraw from our holdout data to get realistic distributions
  of empirical estimates. SEs are estimated based on our planned N of 400.
- We repeat this random drawing process many times.
- We adjust for sampling error/the standard error of the empirical estimate to get the semi-latent accuracy.
- The target quantity is the average standard error of the semi-latent accuracy with which the synthetic estimates predicts the empirical estimates. We also report the maximal standard error across simulations.

We begin with item correlations, then reliabilities, then scale correlations.

```{r warning=F,message=F}
knitr::opts_chunk$set(echo = TRUE, error = T, warning = F, message = F)

# Libraries and Settings

# Libs ---------------------------
library(tidyverse)
library(arrow)
library(glue)
library(psych)
library(lavaan)
library(ggplot2)
library(plotly)
library(gridExtra)
library(semTools)
library(semPlot)

model_name = "ItemSimilarityTraining-20240502-trial12"
#model_name = "item-similarity-20231018-122504"
pretrained_model_name = "all-mpnet-base-v2"

data_path = glue("./")
pretrained_data_path = glue("./")

set.seed(42)

number_of_items <- 300
combinations_items <- choose(number_of_items, 2)
planned_N <- 1200
```


## Precision simulation for synthetic inter-item correlations
```{r}
holdout <- arrow::read_feather(file = file.path(data_path, glue("data/intermediate/{model_name}.raw.osf-bainbridge-2021-s2-0.item_correlations.feather")))

holdout_mapping_data = arrow::read_feather(
  file = file.path(data_path, glue("{model_name}.raw.osf-bainbridge-2021-s2-0.mapping2.feather"))
) %>%
  rename(scale_0 = scale0,
         scale_1 = scale1)

scales <- arrow::read_feather(file.path(data_path, glue("{model_name}.raw.osf-bainbridge-2021-s2-0.scales.feather"))
)

holdout_llm <- holdout %>%
  left_join(holdout_mapping_data %>% select(variable_1 = variable, InstrumentA = instrument, ScaleA = scale_0, SubscaleA = scale_1)) %>%
  left_join(holdout_mapping_data %>% select(variable_2 = variable, InstrumentA = instrument, ScaleA = scale_0, SubscaleA = scale_1))

sim_results <- tibble()
library(lavaan)

for(i in 1:500) {
  items <- holdout %>% select(variable_1) %>% distinct() %>% sample_n(number_of_items) %>% pull(variable_1)

  subset <- holdout %>% filter(variable_1 %in% items, variable_2 %in% items)

  N <- planned_N
  subset <- subset %>% mutate(se = (1 - empirical_r^2)/sqrt(N - 3))
  se2 <- mean(subset$se^2)

  r <- broom::tidy(cor.test(subset$empirical_r, subset$synthetic_r))

  model <- paste0('
    # Latent variables
    PearsonLatent =~ 1*empirical_r

    # Fixing error variances based on known standard errors
    empirical_r ~~ ',se2,'*empirical_r

    # Relationship between latent variables
    PearsonLatent ~~ synthetic_r
  ')

  fit <- sem(model, data = subset)

  sim_results <- bind_rows(sim_results,
    standardizedsolution(fit) %>% filter(lhs == "PearsonLatent", rhs ==  "synthetic_r")
  )
}

sim_results %>% summarise(semi_latent_r = mean(est.std), mean_se = sqrt(mean(se^2)), max_se = max(se)) %>% 
  knitr::kable()
```


