---
title: "Pellert et al. Method: Mean-Centered Embeddings"
output: html_document
date: "`r Sys.Date()`"
---

This document implements the method suggested by Pellert et al. for removing common linguistic features by subtracting the mean embedding vector from each item embedding.

```{r setup, include=FALSE,warning=F,message=F}
knitr::opts_chunk$set(echo = TRUE, error = T, message = F, warning = F)

# Libraries
library(tidyverse)
library(arrow)
library(glue)
library(broom)
library(knitr)

model_name = "ItemSimilarityTraining-20240502-trial12"
pretrained_model_name = "all-mpnet-base-v2"

data_path = glue("./data/intermediate/")

set.seed(42)
source("global_functions.R")
```

## Load Data

```{r load-data}
# Load human data
rr_validation_human_data = arrow::read_feather(
  file = file.path(data_path, glue("{model_name}.raw.validation-study-2024-11-01.human.feather"))
)

# Load mapping data
rr_validation_mapping_data = arrow::read_feather(
  file = file.path(data_path, glue("{model_name}.raw.validation-study-2024-11-01.mapping2.feather"))
)

# Load pre-trained model embeddings
pt_rr_validation_machine_data = rio::import("https://github.com/synth-science/surveybot3000/raw/refs/heads/main/validation_study/embeddings_all-mpnet-base-v2.feather")
pt_rr_validation_machine_data <- pt_rr_validation_machine_data %>% 
  rowwise() %>% 
  mutate(embed_id = list(1:length(embeddings))) %>% 
  unnest(cols = c(embeddings, embed_id)) %>% 
  select(-item) %>% 
  pivot_wider(names_from = id, values_from = embeddings) %>% 
  select(-embed_id)

# Load fine-tuned model embeddings
rr_validation_machine_data = rio::import("https://github.com/synth-science/surveybot3000/raw/refs/heads/main/validation_study/embeddings_surveybot3000.feather")
rr_validation_machine_data <- rr_validation_machine_data %>% 
  rowwise() %>% 
  mutate(embed_id = list(1:length(embeddings))) %>% 
  unnest(cols = c(embeddings, embed_id)) %>% 
  select(-item) %>% 
  pivot_wider(names_from = id, values_from = embeddings) %>% 
  select(-embed_id)

cat("Loaded embeddings:\n")
cat("  Pre-trained dimensions:", nrow(pt_rr_validation_machine_data), "\n")
cat("  Pre-trained items:", ncol(pt_rr_validation_machine_data), "\n")
cat("  Fine-tuned dimensions:", nrow(rr_validation_machine_data), "\n")
cat("  Fine-tuned items:", ncol(rr_validation_machine_data), "\n")
cat("  Human data respondents:", nrow(rr_validation_human_data), "\n")
cat("  Human data items:", ncol(rr_validation_human_data), "\n")
```

## Method: Mean-Centered Embeddings

Following Pellert et al., we compute the mean embedding across all items (the "questionnaire embedding") and subtract it from each item embedding: $\mathbf{y}_i - \bar{\mathbf{y}}$ for item $i$, where $\bar{\mathbf{y}} = \frac{1}{n} \sum_{i=1}^{n} \mathbf{y}_i$.

We implement two variants:
1. **Global centering**: Subtract the mean across all items
2. **Per-instrument centering**: Subtract the mean within each questionnaire/instrument (as Pellert et al. actually did)

```{r center-embeddings}
# Function to center embeddings by subtracting the mean
center_embeddings_global <- function(embedding_matrix) {
  # embedding_matrix: rows = dimensions, columns = items
  # Compute mean across all items (row means)
  mean_embedding <- rowMeans(embedding_matrix, na.rm = TRUE)
  
  # Subtract mean from each item
  centered <- sweep(embedding_matrix, 1, mean_embedding, "-")
  
  return(centered)
}

# Function to center embeddings per instrument
center_embeddings_by_instrument <- function(embedding_matrix, mapping_data) {
  # embedding_matrix: rows = dimensions, columns = items
  # mapping_data: dataframe with 'variable' and 'instrument' columns
  
  centered <- embedding_matrix
  
  # Get unique instruments
  instruments <- unique(mapping_data$instrument)
  
  for (instr in instruments) {
    # Get items belonging to this instrument
    item_ids <- mapping_data %>% 
      filter(instrument == instr) %>% 
      pull(variable)
    
    # Find which columns correspond to these items
    item_cols <- which(colnames(embedding_matrix) %in% item_ids)
    
    if (length(item_cols) > 0) {
      # Compute mean embedding for this instrument
      instr_mean <- rowMeans(embedding_matrix[, item_cols, drop = FALSE], na.rm = TRUE)
      
      # Subtract instrument mean from each item in this instrument
      centered[, item_cols] <- sweep(embedding_matrix[, item_cols, drop = FALSE], 
                                     1, instr_mean, "-")
    }
  }
  
  return(centered)
}

# Global centering
pt_centered_global <- center_embeddings_global(pt_rr_validation_machine_data)
ft_centered_global <- center_embeddings_global(rr_validation_machine_data)

# Per-instrument centering
pt_centered_instrument <- center_embeddings_by_instrument(pt_rr_validation_machine_data, 
                                                          rr_validation_mapping_data)
ft_centered_instrument <- center_embeddings_by_instrument(rr_validation_machine_data,
                                                          rr_validation_mapping_data)

cat("Centered embeddings created\n")
cat("Global centering:\n")
cat("  Pre-trained: mean of means =", mean(rowMeans(pt_centered_global)), "(should be ~0)\n")
cat("  Fine-tuned: mean of means =", mean(rowMeans(ft_centered_global)), "(should be ~0)\n")
cat("\nPer-instrument centering:\n")
cat("  Pre-trained: overall mean =", mean(rowMeans(pt_centered_instrument)), "\n")
cat("  Fine-tuned: overall mean =", mean(rowMeans(ft_centered_instrument)), "\n")
cat("  Number of instruments:", n_distinct(rr_validation_mapping_data$instrument), "\n")
```

## Compute Correlations

Now we compute pairwise correlations (cosine similarities) using the original, globally-centered, and per-instrument-centered embeddings.

```{r compute-correlations}
# Original correlations
pt_original_pairs <- join_pairwise_correlation(rr_validation_human_data, pt_rr_validation_machine_data)
ft_original_pairs <- join_pairwise_correlation(rr_validation_human_data, rr_validation_machine_data)

# Globally mean-centered correlations
pt_centered_global_pairs <- join_pairwise_correlation(rr_validation_human_data, as.data.frame(pt_centered_global))
ft_centered_global_pairs <- join_pairwise_correlation(rr_validation_human_data, as.data.frame(ft_centered_global))

# Per-instrument mean-centered correlations
pt_centered_instrument_pairs <- join_pairwise_correlation(rr_validation_human_data, as.data.frame(pt_centered_instrument))
ft_centered_instrument_pairs <- join_pairwise_correlation(rr_validation_human_data, as.data.frame(ft_centered_instrument))

cat("Computed pairwise correlations:\n")
cat("  Number of item pairs:", nrow(pt_original_pairs), "\n")
```

## Results: Accuracy Comparison

```{r accuracy-comparison}
# Pre-trained model
pt_original_accuracy <- broom::tidy(cor.test(pt_original_pairs$empirical_r, pt_original_pairs$synthetic_r))
pt_centered_global_accuracy <- broom::tidy(cor.test(pt_centered_global_pairs$empirical_r, pt_centered_global_pairs$synthetic_r))
pt_centered_instrument_accuracy <- broom::tidy(cor.test(pt_centered_instrument_pairs$empirical_r, pt_centered_instrument_pairs$synthetic_r))

# Fine-tuned model
ft_original_accuracy <- broom::tidy(cor.test(ft_original_pairs$empirical_r, ft_original_pairs$synthetic_r))
ft_centered_global_accuracy <- broom::tidy(cor.test(ft_centered_global_pairs$empirical_r, ft_centered_global_pairs$synthetic_r))
ft_centered_instrument_accuracy <- broom::tidy(cor.test(ft_centered_instrument_pairs$empirical_r, ft_centered_instrument_pairs$synthetic_r))

# Create comparison table
results <- bind_rows(
  pt_original_accuracy %>% 
    mutate(model = "Pre-trained (SBERT)", method = "Original") %>%
    select(model, method, accuracy = estimate, conf.low, conf.high),
  pt_centered_global_accuracy %>% 
    mutate(model = "Pre-trained (SBERT)", method = "Global mean-centered") %>%
    select(model, method, accuracy = estimate, conf.low, conf.high),
  pt_centered_instrument_accuracy %>% 
    mutate(model = "Pre-trained (SBERT)", method = "Per-instrument mean-centered") %>%
    select(model, method, accuracy = estimate, conf.low, conf.high),
  ft_original_accuracy %>% 
    mutate(model = "Fine-tuned (SurveyBot3000)", method = "Original") %>%
    select(model, method, accuracy = estimate, conf.low, conf.high),
  ft_centered_global_accuracy %>% 
    mutate(model = "Fine-tuned (SurveyBot3000)", method = "Global mean-centered") %>%
    select(model, method, accuracy = estimate, conf.low, conf.high),
  ft_centered_instrument_accuracy %>% 
    mutate(model = "Fine-tuned (SurveyBot3000)", method = "Per-instrument mean-centered") %>%
    select(model, method, accuracy = estimate, conf.low, conf.high)
)

kable(results, digits = 3, caption = "Manifest accuracy (Pearson r) between synthetic and empirical item correlations")
```

## Interpretation

```{r interpretation}
# Calculate improvements
pt_improvement_global <- pt_centered_global_accuracy$estimate - pt_original_accuracy$estimate
pt_improvement_instrument <- pt_centered_instrument_accuracy$estimate - pt_original_accuracy$estimate
ft_improvement_global <- ft_centered_global_accuracy$estimate - ft_original_accuracy$estimate
ft_improvement_instrument <- ft_centered_instrument_accuracy$estimate - ft_original_accuracy$estimate

cat("\n")
cat("Change in accuracy after mean-centering:\n")
cat("\nPre-trained model:\n")
cat("  Global centering:       ", sprintf("%+.3f", pt_improvement_global), "\n")
cat("  Per-instrument centering:", sprintf("%+.3f", pt_improvement_instrument), "\n")
cat("\nFine-tuned model:\n")
cat("  Global centering:       ", sprintf("%+.3f", ft_improvement_global), "\n")
cat("  Per-instrument centering:", sprintf("%+.3f", ft_improvement_instrument), "\n")
cat("\n")

# Determine best method for each model
pt_best <- c("original", "global", "instrument")[which.max(c(pt_original_accuracy$estimate, 
                                                             pt_centered_global_accuracy$estimate, 
                                                             pt_centered_instrument_accuracy$estimate))]
ft_best <- c("original", "global", "instrument")[which.max(c(ft_original_accuracy$estimate, 
                                                             ft_centered_global_accuracy$estimate, 
                                                             ft_centered_instrument_accuracy$estimate))]

cat("Best performing method:\n")
cat("  Pre-trained model: ", pt_best, "\n")
cat("  Fine-tuned model:  ", ft_best, "\n")
```

## Scatter Plots

### Pre-trained Model

```{r pt-scatter, fig.width=15, fig.height=5}
library(patchwork)

p1 <- ggplot(pt_original_pairs, aes(synthetic_r, empirical_r)) +
  geom_abline(linetype = "dashed") +
  geom_point(color = "#00A0B0", alpha = 0.03, size = 1) +
  geom_smooth(method = "lm", color = "#a48500", fill = "#EDC951") +
  annotate("text", x = -1, y = 0.98, hjust = 0, vjust = 1, size = 3,
           label = sprintf("r = %.3f [%.3f, %.3f]", 
                          pt_original_accuracy$estimate,
                          pt_original_accuracy$conf.low,
                          pt_original_accuracy$conf.high)) +
  xlab("Synthetic inter-item correlation") +
  ylab("Empirical") +
  ggtitle("Original") +
  coord_fixed(xlim = c(-1, 1), ylim = c(-1, 1)) +
  theme_bw()

p2 <- ggplot(pt_centered_global_pairs, aes(synthetic_r, empirical_r)) +
  geom_abline(linetype = "dashed") +
  geom_point(color = "#00A0B0", alpha = 0.03, size = 1) +
  geom_smooth(method = "lm", color = "#a48500", fill = "#EDC951") +
  annotate("text", x = -1, y = 0.98, hjust = 0, vjust = 1, size = 3,
           label = sprintf("r = %.3f [%.3f, %.3f]", 
                          pt_centered_global_accuracy$estimate,
                          pt_centered_global_accuracy$conf.low,
                          pt_centered_global_accuracy$conf.high)) +
  xlab("Synthetic inter-item correlation") +
  ylab("Empirical") +
  ggtitle("Global mean-centered") +
  coord_fixed(xlim = c(-1, 1), ylim = c(-1, 1)) +
  theme_bw()

p3 <- ggplot(pt_centered_instrument_pairs, aes(synthetic_r, empirical_r)) +
  geom_abline(linetype = "dashed") +
  geom_point(color = "#00A0B0", alpha = 0.03, size = 1) +
  geom_smooth(method = "lm", color = "#a48500", fill = "#EDC951") +
  annotate("text", x = -1, y = 0.98, hjust = 0, vjust = 1, size = 3,
           label = sprintf("r = %.3f [%.3f, %.3f]", 
                          pt_centered_instrument_accuracy$estimate,
                          pt_centered_instrument_accuracy$conf.low,
                          pt_centered_instrument_accuracy$conf.high)) +
  xlab("Synthetic inter-item correlation") +
  ylab("Empirical") +
  ggtitle("Per-instrument mean-centered") +
  coord_fixed(xlim = c(-1, 1), ylim = c(-1, 1)) +
  theme_bw()

p1 + p2 + p3 + plot_annotation(title = "Pre-trained Model (SBERT)")
```

### Fine-tuned Model

```{r ft-scatter, fig.width=15, fig.height=5}
p4 <- ggplot(ft_original_pairs, aes(synthetic_r, empirical_r)) +
  geom_abline(linetype = "dashed") +
  geom_point(color = "#00A0B0", alpha = 0.03, size = 1) +
  geom_smooth(method = "lm", color = "#a48500", fill = "#EDC951") +
  annotate("text", x = -1, y = 0.98, hjust = 0, vjust = 1, size = 3,
           label = sprintf("r = %.3f [%.3f, %.3f]", 
                          ft_original_accuracy$estimate,
                          ft_original_accuracy$conf.low,
                          ft_original_accuracy$conf.high)) +
  xlab("Synthetic inter-item correlation") +
  ylab("Empirical") +
  ggtitle("Original") +
  coord_fixed(xlim = c(-1, 1), ylim = c(-1, 1)) +
  theme_bw()

p5 <- ggplot(ft_centered_global_pairs, aes(synthetic_r, empirical_r)) +
  geom_abline(linetype = "dashed") +
  geom_point(color = "#00A0B0", alpha = 0.03, size = 1) +
  geom_smooth(method = "lm", color = "#a48500", fill = "#EDC951") +
  annotate("text", x = -1, y = 0.98, hjust = 0, vjust = 1, size = 3,
           label = sprintf("r = %.3f [%.3f, %.3f]", 
                          ft_centered_global_accuracy$estimate,
                          ft_centered_global_accuracy$conf.low,
                          ft_centered_global_accuracy$conf.high)) +
  xlab("Synthetic inter-item correlation") +
  ylab("Empirical") +
  ggtitle("Global mean-centered") +
  coord_fixed(xlim = c(-1, 1), ylim = c(-1, 1)) +
  theme_bw()

p6 <- ggplot(ft_centered_instrument_pairs, aes(synthetic_r, empirical_r)) +
  geom_abline(linetype = "dashed") +
  geom_point(color = "#00A0B0", alpha = 0.03, size = 1) +
  geom_smooth(method = "lm", color = "#a48500", fill = "#EDC951") +
  annotate("text", x = -1, y = 0.98, hjust = 0, vjust = 1, size = 3,
           label = sprintf("r = %.3f [%.3f, %.3f]", 
                          ft_centered_instrument_accuracy$estimate,
                          ft_centered_instrument_accuracy$conf.low,
                          ft_centered_instrument_accuracy$conf.high)) +
  xlab("Synthetic inter-item correlation") +
  ylab("Empirical") +
  ggtitle("Per-instrument mean-centered") +
  coord_fixed(xlim = c(-1, 1), ylim = c(-1, 1)) +
  theme_bw()

p4 + p5 + p6 + plot_annotation(title = "Fine-tuned Model (SurveyBot3000)")
```

## Summary Statistics

```{r summary-stats}
# Summary statistics for synthetic correlations
summary_stats <- bind_rows(
  pt_original_pairs %>% 
    summarise(
      model = "Pre-trained",
      method = "Original",
      mean_synthetic = mean(synthetic_r),
      sd_synthetic = sd(synthetic_r),
      mean_abs_synthetic = mean(abs(synthetic_r))
    ),
  pt_centered_global_pairs %>% 
    summarise(
      model = "Pre-trained",
      method = "Global mean-centered",
      mean_synthetic = mean(synthetic_r),
      sd_synthetic = sd(synthetic_r),
      mean_abs_synthetic = mean(abs(synthetic_r))
    ),
  pt_centered_instrument_pairs %>% 
    summarise(
      model = "Pre-trained",
      method = "Per-instrument mean-centered",
      mean_synthetic = mean(synthetic_r),
      sd_synthetic = sd(synthetic_r),
      mean_abs_synthetic = mean(abs(synthetic_r))
    ),
  ft_original_pairs %>% 
    summarise(
      model = "Fine-tuned",
      method = "Original",
      mean_synthetic = mean(synthetic_r),
      sd_synthetic = sd(synthetic_r),
      mean_abs_synthetic = mean(abs(synthetic_r))
    ),
  ft_centered_global_pairs %>% 
    summarise(
      model = "Fine-tuned",
      method = "Global mean-centered",
      mean_synthetic = mean(synthetic_r),
      sd_synthetic = sd(synthetic_r),
      mean_abs_synthetic = mean(abs(synthetic_r))
    ),
  ft_centered_instrument_pairs %>% 
    summarise(
      model = "Fine-tuned",
      method = "Per-instrument mean-centered",
      mean_synthetic = mean(synthetic_r),
      sd_synthetic = sd(synthetic_r),
      mean_abs_synthetic = mean(abs(synthetic_r))
    )
)

kable(summary_stats, digits = 3, caption = "Summary statistics for synthetic correlations")
```

## Within vs. Across Instrument Analysis

The per-instrument centering likely drops accuracy because it removes valid information about similarities between instruments. Pellert et al. applied this method to a single questionnaire (PVQ-RR), but here we have items from multiple instruments where cross-instrument similarities are meaningful.

Let's examine accuracy separately for within-instrument and across-instrument item pairs:

```{r within-across-analysis}
# Add instrument information to each pair
add_instrument_info <- function(pairs, mapping) {
  pairs %>%
    left_join(mapping %>% select(variable_1 = variable, instrument_1 = instrument), 
              by = "variable_1") %>%
    left_join(mapping %>% select(variable_2 = variable, instrument_2 = instrument), 
              by = "variable_2") %>%
    mutate(same_instrument = instrument_1 == instrument_2)
}

# Add instrument info to all datasets
pt_original_pairs_instr <- add_instrument_info(pt_original_pairs, rr_validation_mapping_data)
pt_centered_global_pairs_instr <- add_instrument_info(pt_centered_global_pairs, rr_validation_mapping_data)
pt_centered_instrument_pairs_instr <- add_instrument_info(pt_centered_instrument_pairs, rr_validation_mapping_data)

ft_original_pairs_instr <- add_instrument_info(ft_original_pairs, rr_validation_mapping_data)
ft_centered_global_pairs_instr <- add_instrument_info(ft_centered_global_pairs, rr_validation_mapping_data)
ft_centered_instrument_pairs_instr <- add_instrument_info(ft_centered_instrument_pairs, rr_validation_mapping_data)

# Calculate accuracy within and across instruments
accuracy_by_context <- bind_rows(
  # Pre-trained
  pt_original_pairs_instr %>% group_by(same_instrument) %>%
    summarise(broom::tidy(cor.test(empirical_r, synthetic_r)), n_pairs = n()) %>%
    mutate(model = "Pre-trained", method = "Original"),
  pt_centered_global_pairs_instr %>% group_by(same_instrument) %>%
    summarise(broom::tidy(cor.test(empirical_r, synthetic_r)), n_pairs = n()) %>%
    mutate(model = "Pre-trained", method = "Global mean-centered"),
  pt_centered_instrument_pairs_instr %>% group_by(same_instrument) %>%
    summarise(broom::tidy(cor.test(empirical_r, synthetic_r)), n_pairs = n()) %>%
    mutate(model = "Pre-trained", method = "Per-instrument mean-centered"),
  
  # Fine-tuned
  ft_original_pairs_instr %>% group_by(same_instrument) %>%
    summarise(broom::tidy(cor.test(empirical_r, synthetic_r)), n_pairs = n()) %>%
    mutate(model = "Fine-tuned", method = "Original"),
  ft_centered_global_pairs_instr %>% group_by(same_instrument) %>%
    summarise(broom::tidy(cor.test(empirical_r, synthetic_r)), n_pairs = n()) %>%
    mutate(model = "Fine-tuned", method = "Global mean-centered"),
  ft_centered_instrument_pairs_instr %>% group_by(same_instrument) %>%
    summarise(broom::tidy(cor.test(empirical_r, synthetic_r)), n_pairs = n()) %>%
    mutate(model = "Fine-tuned", method = "Per-instrument mean-centered")
) %>%
  select(model, method, same_instrument, accuracy = estimate, conf.low, conf.high, n_pairs)

# Format for display
accuracy_by_context %>%
  mutate(context = if_else(same_instrument, "Within instrument", "Across instruments"),
         accuracy_ci = sprintf("%.3f [%.3f, %.3f]", accuracy, conf.low, conf.high)) %>%
  select(model, method, context, accuracy_ci, n_pairs) %>%
  kable(caption = "Accuracy for within-instrument vs. across-instrument item pairs")
```

### Interpretation

**Per-instrument mean-centering damages accuracy both within AND across instruments**, suggesting that what Pellert et al. characterized as "common linguistic features" and "unneeded information" actually contains valid psychological signal.

Key findings:

1. **Across-instrument pairs**: Per-instrument centering removes meaningful semantic differences between instruments (e.g., depression items vs. personality items have genuinely different content)

2. **Within-instrument pairs**: Even here, per-instrument centering reduces accuracy, suggesting the "questionnaire embedding" (mean vector) captures valid shared content rather than mere linguistic artifacts

**Why might the instrument mean contain signal rather than noise?**

- Items within a questionnaire share substantive psychological content (e.g., all depression items reference negative affect)
- This shared content creates meaningful correlations that the model should detect
- Removing it leaves only item-specific deviations, discarding valid information about how items relate to the construct

**Implications**: The Pellert et al. approach may be problematic even for single-questionnaire applications if the questionnaire has coherent substantive content. The "high-dimensional linguistic features" they remove may actually be capturing real psychological similarities.

Since our validation study includes `r round(mean(!pt_original_pairs_instr$same_instrument)*100, 1)`% cross-instrument pairs, the overall accuracy drops substantially, but the within-instrument degradation is equally concerning.
```

## Summary of Findings 

Pellert et al. (2024) propose "Survey and Questionnaire Item Embeddings Differentials" (SQuID), a method for improving the quality of semantic embeddings for psychometric items. Their approach involves subtracting the mean embedding vector across all questionnaire items from each individual item embedding. The authors frame this operation as removing "common linguistic features of natural language use" that artificially inflate similarities between items and obscure meaningful psychological differences. They report substantial improvements in their validation study using the 57-item Portrait Values Questionnaire-Revised (PVQ-RR), particularly for pre-trained models.

We tested whether this method generalizes to our validation dataset, which contains 246 items from multiple psychological instruments. We implemented two variants of the method: global mean-centering (subtracting the mean across all 246 items) and per-instrument mean-centering (subtracting the instrument-specific mean from items within each questionnaire, as Pellert et al. actually did with their single-questionnaire dataset).

The empirical results reveal a striking pattern. For the pre-trained SBERT model, global mean-centering produces a modest improvement in accuracy from 0.302 to 0.340 (but the graph still looks croissant shaped, so it does not solve the problem with negative correlations). However, for the fine-tuned model, mean-centering consistently degrades performance. Global mean-centering reduces accuracy from 0.570 to 0.551, while per-instrument mean-centering reduces it further to 0.295. Most critically, when examining item pairs within the same instrument, per-instrument mean-centering reduces accuracy from approximately 0.54 to 0.38. This within-instrument degradation directly contradicts the interpretation that the instrument-level mean embedding represents removable linguistic noise.

The theoretical issue becomes clear when considering what items within a well-designed psychological instrument actually share. Items measuring depression, for example, reference related psychological states such as sadness, anhedonia, fatigue, and worthlessness. These items should correlate because they measure facets of the same underlying construct. Their embeddings are genuinely similar because they describe related psychological content. The instrument-level mean embedding captures this shared construct-level semantic content, not merely superficial linguistic features.

When the instrument mean is subtracted from each item embedding, the operation removes the semantic core that defines what the instrument measures. What remains are only item-specific deviations from the construct mean. While this might preserve some information about how items differ in their particular instantiation of the construct, it destroys information about how strongly each item relates to the overall construct. For a fine-tuned model that has learned to encode psychological constructs efficiently, this removal of shared construct information directly damages the model's ability to predict empirical item correlations.

The differential effects across models provide additional insight. The pre-trained SBERT model benefits slightly from global mean-centering, likely because it does contain substantial generic linguistic patterns that happen to be consistent across survey items but unrelated to psychological content. The fine-tuned model, having learned to use the embedding space specifically for psychological similarity, encodes less generic linguistic noise and more construct-specific information in its shared embeddings. Removing the mean therefore damages the fine-tuned model more severely.

These findings suggest that Pellert et al.'s success may not reflect a general principle about removing linguistic artifacts from psychological embeddings. Instead, their method may have been effective for their specific combination of a pre-trained model and a single, highly coherent questionnaire. The method fails when applied to a fine-tuned model or to a diverse collection of instruments, precisely because it removes valid psychological signal rather than mere linguistic noise. Researchers should be cautious about applying this preprocessing step, particularly when working with models that have been trained on psychological content or when analyzing multiple instruments where construct-level differences contain meaningful information.
